Remote Collection of Respiratory Samples (NACFC 2023)

### 2023-03-31 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd
added Cohen's kappa

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
  )
# Shorten organism names
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
```

@@

# Sensitivity and specificity

```{r}
sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
  t <- table(other_test, gold_standard)
  if (all(dim(t) == c(1, 1))) {
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    cat("All samples marked as", unique(gold_standard))
    cat("\n")
  } else {
    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))

    dimnames(t) <- lapply(dimnames(t), function(n) {
      n <- sub("Checked", "Present", n)
      n <- sub("Unchecked", "Absent", n)
    })
    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    print(knitr::kable((as.data.frame.matrix(t))))
    cat("\n")
    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", round(spec, 3))
    cat("\n")
  }
}
```

@@

>>>>

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(irr)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
  )
# Shorten organism names
bugs = colnames(df)[grep("If.culture.was.positive..select.all.organisms.isolated...choice.",colnames(df))]
bugs <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", bugs)
bugs = bugs[1:16]
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
```

@@

# Pairwise Comparisons

Cohen's kappa is an agreement statistic ranging from -1 (perfect disagreement) to 1 (perfect agreement). A rough guideline for interpreting Cohen's kappa is:

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

```{r}
sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
  t <- table(other_test, gold_standard)
  if (all(dim(t) == c(1, 1))) {
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("All samples marked as", unique(gold_standard))
    cat("\n")
  } else {
    # Sensitivity and specificity
    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))
    # Kappa
    k = kappa2(cbind(gold_standard,other_test))
    # Formatting
    dimnames(t) <- lapply(dimnames(t), function(n) {
      n <- sub("Checked", "Present", n)
      n <- sub("Unchecked", "Absent", n)
    })
    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    print(knitr::kable((as.data.frame.matrix(t))))
    cat("\n")
    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", paste0(round(spec, 3),"."))
    cat("\n")
    cat("Cohen's kappa for the two tests was", round(k$value, 3), paste0("(p = ", format.pval(k$p.value,digits = 3, eps = 0.001),")"))
    cat("\n")
  }
}
```

@@

## Delayed Processing

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
  df$Methicillin.sensitive.Staph.aureus..2,
  type = "Delayed Processing", bug = "MSSA"
)
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
  df$Methicillin.resistant.Staph.aureus..2,
  type = "Delayed Processing", bug = "MRSA"
)
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..2,
  type = "Delayed Processing", bug = "A. xylosoxidans"
)
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..2,
  type = "Delayed Processing", bug = "Achromobacter (not xylosoxidans)"
)
```

```{r results='asis'}
sens_spec(df$Acinetobacter., df$Acinetobacter..2,
  type = "Delayed Processing", bug = "Acinetobacter"
)
```

```{r results='asis'}
sens_spec(df$H..influenzae., df$H..influenzae..2,
  type = "Delayed Processing", bug = "H. influenzae"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa., df$P..aeruginosa..2,
  type = "Delayed Processing", bug = "P. aeruginosa"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...2,
  type = "Delayed Processing", bug = "P. aeruginosa (mucoid)"
)
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..2,
  type = "Delayed Processing", bug = "Pseudomonas (not aeruginosa)"
)
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..2,
  type = "Delayed Processing", bug = "B. cepacia complex"
)
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..2,
  type = "Delayed Processing", bug = "Stenotrophomonas"
)
```

```{r results='asis'}
sens_spec(df$Aspergillus., df$Aspergillus..2,
  type = "Delayed Processing", bug = "Aspergillus"
)
```

```{r results='asis'}
sens_spec(df$Penicillium.species., df$Penicillium.species..2,
  type = "Delayed Processing", bug = "Penicillium species"
)
```

```{r results='asis'}
sens_spec(df$Sceosporium.species., df$Sceosporium.species..2,
  type = "Delayed Processing", bug = "Sceosporium species"
)
```

```{r results='asis'}
sens_spec(df$M..avium.complex., df$M..avium.complex..2,
  type = "Delayed Processing", bug = "M. avium complex"
)
```

```{r results='asis'}
sens_spec(df$M..abscessus., df$M..abscessus..2,
  type = "Delayed Processing", bug = "M. abscessus"
)
```

# Agreement between all 3 methods

```{r}

```
